Cross Validation Approaches for Penalized Cox Regression

### 2018-05-26 main.tex
Edits to Ch. 2 intro

In the Cox model, the hazard function for subject $i$ is given by 
\begin{equation}
  h_{i}(t) = h_{0}(t) \text{exp}( X_{i}^{T} \beta),
\end{equation} 
where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for subject $i$ with covariates $X_i$ relative to that baseline.  The estimation of the coefficients is obtained by maximizing a partial likelihood.  Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
\begin{equation}
  l_{i}(\beta) = \left \{\frac{\text{exp} ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\text{exp} ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
\end{equation}
where $R(t_{i})$ denotes the set of observations at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
\begin{equation}
  L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta).
\end{equation}
%% \begin{equation}
%%   L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta) = \prod_{j=1}^{m} \frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}.
%% \end{equation}

Penalized Cox Regression is an extension of the regular cox regression model. Estimates of the $\beta$ coefficients are obtained by minimizing the objective function
\begin{equation}
  Q(\beta |X, y) = - \frac{1}{n}  \sum_{j=1}^{m} \text{log} \left \{\frac{\text{exp} ( X_{j}^{T} \beta)}{\sum_{ k \in R(t_{j})}\text{exp} ( X_{k}^{T} \beta)}) \right \} + P_{\lambda}(\beta).
\end{equation}
The first part of the objective function is based on the partial likelihood. The second part is a penalty term that depends on some regularization parameters. In this paper, only the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$ was considered. The LASSO cox regression can be used to build prediction models for survival outcome based on high dimensional data, where the number of covariates p $\gg$ n. An example of its application would be to make predictions of cancer patients' the overall survival time with their certain gene expressions.

Selecting the tuning parameter $\lambda$ is crucial to getting LASSO estimations for $\beta$. The LASSO penalty enforces sparcity to the regression model, so that it carries out an automatic variable selection along the solution path. As $\lambda$ increases, more coefficients would be set to 0 and fewer variables will be selected by the model. When $\lambda$ is too small, the model tends to be redundant. When $\lambda$ is too large, the model would be too conservative to select any variables. In either case, model's predictive accuracy would be affected.

>>>>

In the Cox model, the hazard function for subject $i$ is given by 
\begin{equation}
  h_{i}(t) = h_{0}(t) \exp( X_{i}^{T} \beta),
\end{equation} 
where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for a subject with covariates $X_i$ relative to that baseline.  The estimation of the coefficients is obtained by maximizing a partial likelihood.  Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
\begin{equation}
  l_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
\end{equation}
where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
\begin{equation}
  L(\beta) =\prod_{i = 1}^{n} l_{i}(\beta).
\end{equation}

Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
\begin{equation}
  Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
\end{equation}
where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$. In this paper, we focus upon the LASSO penalty $P_{\lambda}(\beta) = \lambda (\sum_{j} |\beta_{j}|)$, although the methods we propose can be used with any penalty (indeed, can also be used for model selection in unpenalized Cox regression).  LASSO cox regression is particularly useful for high dimensional data where the number of covariates p $\gg$ n; for example, in the prediction of cancer patients' overall survival based on genome-wide expression measurements.

Selecting $\lambda$ is critical to LASSO estimation. LASSO estimates are sparse, in the sense that some coefficients are estimated to be exactly zero, thereby carrying out automatic variable selection along the solution path. At large values of $\lambda$, most or all of the coefficients are 0, but as $\lam$ decreases, more covariates are selected into the model. If $\lam$ is too large or too small, the model's predictive accuracy suffers.
